---
title: "Experiments"
author: "Raphael Morsomme"
date: "`r Sys.Date()`"
output: html_document
---

This file contains the code necessary to run the simulations presented in the manuscript. Running all experiments with the settings used in the manuscript takes around 12 hours, with *Experiment 3* taking the longest (10 hours). To reduce the running time, one can run a single simulation or decreasing the number of MCMC draws from 1 million to 100,000.

## Installation of R package PDSIR

```{r package}
devtools::install_github("rmorsomme/PDSIR")
library(PDSIR)
```

## Creation of paths for figures and mcmc draws for the various experiments

```{r}
path_figure <- "Output/Figures/"
path_data   <- "Output/RDATA/"

# Experiment 1
path_figure_E1 <- paste0(path_figure, "E1")
path_data_E1   <- paste0(path_data  , "E1")

# Experiment 2
path_figure_E2 <- paste0(path_figure, "E2")

# Experiment 3
path_figure_E3 <- paste0(path_figure, "E3")
path_data_E3   <- paste0(path_data  , "E3")

# Experiment 4
path_data_E4   <- paste0(path_data  , "E4")

# Experiment 5
path_figure_E5 <- paste0(path_figure, "E5")
path_data_E5   <- paste0(path_data  , "E5")

# Experiment 6
path_figure_E6_joint  <- paste0(path_figure, "E6/joint")
path_figure_E6_single <- paste0(path_figure, "E6/single")
path_data_E6          <- paste0(path_data  , "E6")

# Miscellaneous
path_figure_misc <- paste0(path_figure, "misc")
```



## Simulation studies

### Experiments 1 -- proof of concept

```{r}
set.seed(0)
output_E1 <- experiment_1_proof_of_concept(
  N = 1e6, iota_dist = "weibull", theta_0_factor = 0.1,
  path = path_figure_E1, plot_id = "E1"
  )

summary_E1_short <- experiment_1_output_analysis(
  output_E1, iota_dist = "weibull",
  plot_id = "E1_short", path = path_figure_E1,
  burnin = 1e3, n_max = 1e4
)
summary_E1 <- experiment_1_output_analysis(
  output_E1, iota_dist = "weibull",
  plot_id = "E1", path = path_figure_E1,
  burnin = 1e4
)

#
# Output of interest
print(output_E1$Y$T_k)                   # observed infection counts
print(output_E1$MC$run_time)             # running time
print(output_E1$MC$rate_accept)          # M-H acceptance rate
print(summary_E1$summary_burn$ESS)       # effective sample size
print(summary_E1$summary_burn$post_stat) # summary statistics of posterior draws

save(output_E1, file = paste0(path_data_E1, "/mcmc.RDATA"))
```


### Experiment 2 -- comparison of the SIR and PD-SIR processes

```{r}
set.seed(0)
experiment_2_PDSIR_trajectories(path = path_figure_E2)
```


### Experiment 3 -- impact of rho on mixing

```{r}
set.seed(0)
output_E3 <- experiment_3_acceptance_vs_rho(
  N = 1e6, iota_dist = "weibull",
  S0s = c(250, 500, 1000), R0s = c(2.2, 2.5, 3)
  )

experiment_3_output_analysis(
  output_E3, path = path_figure_E3
  )

save(output_E3, file = paste0(path_data_E3, "/mcmc.RDATA"))
```


### Experiment 4 -- frequentist coverage

The following code chunk is designed to be run on Slurm as an array with, say, 200 tasks. Its output is already present in the directory `r path_data_E4` for simplicity. Note that the code can easily be modified to be run as a for-loop in R. The second code chunk processes the result of the 2000 simulations (200 tasks with m=10 simulations each).

```{r, eval=FALSE}
a <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID")) # index value for array on Slurm
output_E4 <- experiment_4_coverage(
  N = 1e6, m = 10, thin = 1, a = a
)
save(output_E4, file = paste0(path_data_E4, "/E4_", a, ".RDATA"))
```



```{r}
load("Output/RDATA/E4/E4_1.RDATA")
results <- output_E4
for(a in 2 : 200) { # 200 realizations of the previous code chunk
  load(paste0(path_data_E4, "/E4_", a, ".RDATA"))
  results <- rbind(results, output_E4)
}

summary_E4 <- experiment_4_output_analysis(results, path = path_figure_E4)
print(summary_E4, 2)
```


### Experiment 6 -- comparison of the block and single-site samplers

```{r}
set.seed(0)
output_E6_single <- experiment_6_single_site_update(
  N = 1e6, save_fig = FALSE, thin = 10, iota_dist = "weibull"
  )

set.seed(0) # use same seed to ensure that same observed data are used
output_E6_joint  <- experiment_1_proof_of_concept(
  N = 1e6, save_fig = FALSE, thin = 10, rho = 1/10, iota_dist = "weibull"
  )

save(output_E6_single, file = paste0(path_data_E6, "/mcmc_single.RDATA"))
save(output_E6_joint, file = paste0(path_data_E6, "/mcmc_joint.RDATA"))
```

## Case study -- Ebola 2013-2016

```{r}
set.seed(0)
output_E5 <- experiment_5_ebola(
  N = 1e6, rho = 0.2, thin = 10,
  iota_dist = "weibull",
  path_data = "Input/Ebola/Guinea_modified.csv"
  )
save(output_E5, file = paste0(path_data_E5, "/mcmc.RDATA"))

print(output_E5$MC$run_time) # run time
output_E5$MC$rate_accept     # M-H acceptance rate
```

## Miscellaneous

```{r}
draw_mu   (path = path_figure_misc)
draw_gamma(path = path_figure_misc)
```

